---
title: "Week11-AE-RMarkdown"
output:
  pdf_document: default
  html_document: default
date: "2024-04-25"
---

```{r}
library(tidyverse)
library(scales)
library(janitor)
library(colorspace)

#install.packages('gganimate')
library(gganimate)

aqi_levels <- tribble(
  ~aqi_min, ~aqi_max, ~color,    ~level,
  0,        50,       "#D8EEDA", "Good",
  51,       100,      "#F1E7D4", "Moderate",
  101,      150,      "#F8E4D8", "Unhealthy for sensitive groups",
  151,      200,      "#FEE2E1", "Unhealthy",
  201,      300,      "#F4E3F7", "Very unhealthy",
  301,      400,      "#F9D0D4", "Hazardous"
)


# Data is in a wide format
# pr_* - political rights, scores 1-7 (1 highest degree of freedom, 7 the lowest)
# cl_* - civil liberties, scores 1-7 (1 highest degree of freedom, 7 the lowest)
# status_* - freedom status (Free, Partly Free, and Not Free)

freedom <- read_csv(
  file = "freedom.csv",
  na = "-"
)

syr_2023 <- read_csv(file = "ad_aqi_tracker_data-2023.csv")

syr_2023 <- syr_2023 |>
  clean_names() |>
  mutate(date = mdy(date))
```

```{r}
syr_2023 |>
  ggplot(aes(x = date, y = aqi_value, group = 1)) +
  # plot the AQI in Syracuse
  geom_line(linewidth = 1, alpha = 0.5)
```

# Task 1. Add color shading to the plot based on the AQI guide. The color palette does not need to match the specific colors in the table yet.
```{r}
syr_plot <- syr_2023 |>
  ggplot(aes(x = date, y = aqi_value, group = 1)) +
  # shade in background with colors based on AQI guide
  geom_rect(
    data = ...,
    aes(
      ymin = ..., ymax = ...,
      xmin = ..., xmax = ...,
      x = ..., y = ..., fill = color
    )
  ) +
  # plot the AQI in Syracuse
  geom_line(linewidth = 1, alpha = 0.5)

syr_plot
```

# Task 2. Use the hexidecimal colors from the dataset for the color palette.

```{r}
syr_plot_colored <- syr_plot +
  ...()

syr_plot_colored
```

# Task 3. Label each AQI category on the chart
Incorporate text labels for each AQI value directly into the graph. To accomplish this, you need to:

- Calculate the midpoint AQI value for each category
- Add a geom_text() layer to the plot with the AQI values positioned at the midpoint on the y-axis

Extend the range of the x-axis to provide more horizontal space for the AQI category labels without interfering with the trend line.

```{r}
aqi_levels <- aqi_levels |>
  mutate(aqi_mid = ((aqi_min + aqi_max) / 2))

syr_plot_labelled <- syr_plot_colored +
  scale_x_date(
    name = NULL, date_labels = "%b %Y",
    limits = c(ymd("2023-01-01"), ymd("2024-03-01"))
  ) +
  # add text labels for each AQI category
  geom_text(
    data = ...,
    aes(x = ..., y = aqi_mid, label = level),
    hjust = 1, size = 6, fontface = "bold", color = "white",
    family = "Atkinson Hyperlegible"
  )

syr_plot_labelled
```

# Task 4. Clean up the plot.

Add a meaningful title, axis labels, caption, etc.
```{r}

syr_plot_labelled <- syr_plot_labelled +
  theme_minimal() +
  ...()

syr_plot_labelled
```

---

From this task we will be using the `freedom.csv` data.

```{r}
freedom_to_plot <- freedom |>
  # calculate rowwise standard deviations (one row per country)
  rowwise() |>
  mutate(sd = sd(c_across(contains("cl_")), na.rm = TRUE)) |>
  ungroup() |>
  # find the 15 countries with the highest standard deviations
  relocate(country, sd) |>
  slice_max(order_by = sd, n = 15) |>
  # only keep countries with complete observations - necessary for future plotting
  drop_na()
freedom_to_plot
```

```{r}
# calculate position rankings rather than raw scores
freedom_ranked <- freedom_to_plot |>
  # only keep columns with civil liberties scores
  select(country, contains("cl_")) |>
  # wrangle the data to a long format
  pivot_longer(
    cols = -country,
    names_to = "year",
    values_to = "civil_liberty",
    names_prefix = "cl_",
    names_transform = list(year = as.numeric)
  ) |>
  # calculate rank within year - larger is worse, so reverse in the ranking
  group_by(year) |>
  mutate(rank_in_year = rank(-civil_liberty, ties.method = "first")) |>
  ungroup() |>
  # highlight Venezuela
  mutate(is_venezuela = if_else(country == "Venezuela", TRUE, FALSE))
freedom_ranked
```

# Task 5. Fill in the code below to create a faceted bar plot by year.

```{r}
freedom_faceted_plot <- freedom_ranked |>
  # civil liberty vs freedom rank
  ggplot(aes(x = ..., y = ..., fill = ...)) +
  geom_col(show.legend = FALSE) +
  # change the color palette for emphasis of Venezuela
  scale_fill_manual(values = c("gray", "red")) +
  # facet by year
  facet_wrap(vars(year)) +
  # create explicit labels for civil liberties score,
  # leaving room for country text labels
  scale_x_continuous(
    limits = ...,
    breaks = ...
  ) +
  geom_text(
    hjust = "...",
    aes(label = ...),
    x = ...
  ) +
  # remove extraneous theme/label components
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_blank()
  ) +
  labs(x = NULL, y = NULL)

freedom_faceted_plot
```
# Task 6. Fill in the code below to turn the facet plot into an animation showing the data by year.

```{r}
# If your code generates a bunch of png's instead of showing an animated object:
# Install this:
#install.packages("magick")
# And restart your R session
```

```{r}
freedom_bar_race <- freedom_faceted_plot +
  # remove faceting
  facet_null() +
  # label the current year in the top corner of the plot
  geom_text(
    x = ..., y = ...,
    hjust = "...",
    aes(label = ...),
    size = ...
  ) +
  # define group structure for transitions
  aes(group = ...) +
  # temporal transition - ensure integer value for labeling
  transition_time(...) +
  labs(
    title = "Civil liberties rating, {frame_time}",
    subtitle = "1: Highest degree of freedom - 7: Lowest degree of freedom"
  )

# basic transition
animate(freedom_bar_race, nframes = 30, fps = 2)
```

```{r}
# smoother transition - might take a while to render
animate(freedom_bar_race, nframes = 300, fps = 100, start_pause = 10, end_pause = 10)
```

