# Plotly

```{r}
library(gapminder)

#install.packages('plotly')
library(plotly)

gapminder_2007 <- filter(
  gapminder,
  year == 2007
)

my_plot <- ggplot(
  data = gapminder_2007,
  mapping = aes(
    x = gdpPercap, y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  scale_x_log10() +
  theme_minimal()

ggplotly(my_plot)
```

# Plotly tooltips
```{r}
my_plot <- ggplot(
  data = gapminder_2007,
  mapping = aes(
    x = gdpPercap, y = lifeExp,
    color = continent
  )
) +
  geom_point(aes(text = country)) +
  scale_x_log10() +
  theme_minimal()

ggplotly(
  my_plot,
  tooltip = "text"
)
```


# Plotly with bars
```{r}
car_hist <- ggplot(
  data = mpg,
  mapping = aes(x = hwy)
) +
  geom_histogram(
    binwidth = 2,
    boundary = 0,
    color = "white"
  )

ggplotly(car_hist)
```

# Creating interactive World Bank plot
```{r}
library(tidyverse)
library(WDI)
library(scales)
library(plotly)
library(colorspace)
#install.packages('ggbeeswarm')
library(ggbeeswarm)

# get World Bank indicators
indicators <- c(
  population = "SP.POP.TOTL",
  prop_women_parl = "SG.GEN.PARL.ZS",
  gdp_per_cap = "NY.GDP.PCAP.KD"
)

wdi_parl_raw <- WDI(
  country = "all", indicators, extra = TRUE,
  start = 2022, end = 2022
)

# keep actual "economies" (this may take a while to run)
wdi_clean <- wdi_parl_raw |>
  filter(region != "Aggregates")
glimpse(wdi_clean)
```

```{r}
wdi_2022 <- wdi_clean |>
  filter(year == 2022) |>
  drop_na(prop_women_parl) |>
  # Scale this down from 0-100 to 0-1 so that scales::label_percent() can format
  # it as an actual percent
  mutate(prop_women_parl = prop_women_parl / 100)

static_plot <- ggplot(
  data = wdi_2022,
  mapping = aes(y = fct_rev(region), x = prop_women_parl, color = region)
) +
  geom_quasirandom() +
  scale_x_continuous(labels = label_percent()) +
  scale_color_discrete_qualitative(guide = "none") +
  labs(x = "% women in parliament", y = NULL, caption = "Source: The World Bank") +
  theme_bw(base_size = 14)
static_plot
```
Make it interactive

```{r}
ggplotly(static_plot)
```

Modifying the tooltip

```{r}
static_plot_tooltip <- ggplot(
  data = wdi_2022,
  mapping = aes(y = fct_rev(region), x = prop_women_parl, color = region)
) +
  geom_quasirandom(
    mapping = aes(text = country)
  ) +
  scale_x_continuous(labels = label_percent()) +
  scale_color_discrete_qualitative() +
  labs(x = "% women in parliament", y = NULL, caption = "Source: The World Bank") +
  theme_bw(base_size = 14) +
  theme(legend.position = "none")

ggplotly(static_plot_tooltip,
  tooltip = "text"
)
```


Creating custom tooltips with the format

```
Name of country
X% women in parliament
```

Generate a new column using `mutate()` with the required character string

- `str_glue()` to combine character strings with data values
- The `<br>` is HTML code for a line break
- Use the `label_percent()` function to format numbers as percents

```{r}
wdi_2022 <- wdi_clean |>
  filter(year == 2022) |>
  drop_na(prop_women_parl) |>
  mutate(
    prop_women_parl = prop_women_parl / 100,
    fancy_label = str_glue("{country}<br>{label_percent(accuracy = 0.1)(prop_women_parl)} women in parliament")
  )
wdi_2022 |>
  select(country, prop_women_parl, fancy_label) |>
  head()
```


```{r}
static_plot_tooltip_fancy <- ggplot(
  data = wdi_2022,
  mapping = aes(y = fct_rev(region), x = prop_women_parl, color = region)
) +
  geom_quasirandom(
    mapping = aes(text = fancy_label)
  ) +
  scale_x_continuous(labels = label_percent()) +
  scale_color_discrete_qualitative() +
  labs(x = "% women in parliament", y = NULL, caption = "Source: The World Bank") +
  theme_bw(base_size = 14) +
  theme(legend.position = "none")

ggplotly(static_plot_tooltip_fancy,
  tooltip = "text"
)
```