---
title: "Week9-AE-RMarkdown"
output: html_document
date: "2024-03-27"
---

```{r}
library(tidyverse)
library(readxl)
library(scales)
#install.packages('janitor')
library(janitor)
```

```{r}
nurses <- read_csv("nurses.csv") |> clean_names()

# subset to three states
nurses_subset <- nurses |>
  filter(state %in% c("California", "New York", "North Carolina"))
```

The following code chunk demonstrates how to add alternative text to a bar chart. The alternative text is added to the chunk header using the `fig-alt` chunk option. The text is written in Markdown and can be as long as needed. Note that fig-cap is not the same as `fig-alt`.

```{r}
#| label: nurses-bar
#| fig-cap: "Total employed Registered Nurses"
#| fig-alt: "The figure is a bar chart titled 'Total employed Registered
#| Nurses' that displays the numbers of registered nurses in three states
#| (California, New York, and North Carolina) over a 20 year period, with data
#| recorded in three time points (2000, 2010, and 2020). In each state, the
#| numbers of registered nurses increase over time. The following numbers are
#| all approximate. California started off with 200K registered nurses in 2000,
#| 240K in 2010, and 300K in 2020. New York had 150K in 2000, 160K in 2010, and
#| 170K in 2020. Finally North Carolina had 60K in 2000, 90K in 2010, and 100K
#| in 2020."

nurses_subset |>
  filter(year %in% c(2000, 2010, 2020)) |>
  ggplot(aes(x = state, y = total_employed_rn, fill = factor(year))) +
  geom_col(position = "dodge") +
  scale_fill_viridis_d(option = "E") +
  scale_y_continuous(labels = label_number(scale = 1/1000, suffix = "K")) +
  labs(
    x = "State", y = "Number of Registered Nurses", fill = "Year",
    title = "Total employed Registered Nurses"
  ) +
  theme(
    legend.background = element_rect(fill = "white", color = "white"),
    legend.position = c(0.85, 0.75)
    )
```

# Task 1. Add alt text to line chart
```{r}
# Your label here
# Your caption here
# Your alt text here

nurses_subset |>
  ggplot(aes(x = year, y = annual_salary_median, color = state)) +
  geom_line() +
  scale_y_continuous(labels = label_dollar(scale = 1/1000, suffix = "K")) +
  labs(
    x = "Year", y = "Annual median salary", color = "State",
    title = "Annual median salary of Registered Nurses"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(0.1, 0.9, 0.1, 0.1, "in")
    )
```

# Task 2. Direct labelling instead of legends

Create a version of the same line chart but using direct labels instead of legends. 

You can achieve this with just `geom_text()` but you can also check out https://r-graph-gallery.com/web-line-chart-with-labels-at-end-of-line.html for a fancier way of achieving this.

```{r}
# YOUR CODE HERE

```

# Task 3. Colorblind-friendly plots
Use `colorblindr` for colorblind-friendly palettes.
```{r}
#remotes::install_github("wilkelab/cowplot")
#install.packages("colorspace", repos = "http://R-Forge.R-project.org")
#remotes::install_github("clauswilke/colorblindr")
library(colorblindr)
```

Try out colorblind simulations at http://hclwizard.org/cvdemulator/ or |> your plot to `cvd_grid()` to see the plot in various color-vision-deficiency simulations.

With the line chart from Task 1, create 3 different plots: one with the default color scale, one with the `viridis` color scale, and one with the `OkabeIto` color scale from `colorblindr`. Show the `cvd_grid()` of each plot and describe the simulated effectiveness of the color scales for colorblind viewers.

```{r}
# YOUR CODE HERE for default color scale

```

What do you think of the default color scale effectiveness for colorblind viewers?

YOUR ANSWER HERE

---

```{r}
# YOUR CODE HERE for viridis color scale

```

What do you think of the viridis color scale effectiveness for colorblind viewers?

YOUR ANSWER HERE

---

```{r}
# YOUR CODE HERE for OkabeIto color scale

```

What do you think of the OkabeIto color scale effectiveness for colorblind viewers? 

YOUR ANSWER HERE


