---
title: "COMP4010 - Week 1"
output:
  html_document: default
  pdf_document: default
date: "2024-02-20"
---

# Week 1

# Application Exercises

## Task 1.

-   Data:
-   Mapping:
-   Statistical transformation:
-   Geometric object:
-   Position adjustment:
-   Scale:
-   Coordinate system:
-   Faceting:

## Task 2.

```{r}
# Your code here

```

## Task 3.

```{r}
# Your code here

```

# Reading Material

## Hello World! but in R

Create a chunk below and create a vector of numbers and calculate its mean. (This is akin to printing 'Hello World' in other languages, statisticians are not as fun :D)

```{r}
myVector <- c(1,2,3,4)
mean(myVector)
```

## Using CRAN to install ggplot2

Alternatively, you can just run this in the console below.

```{r}
#install.packages("ggplot2") # uncomment to install
library(ggplot2) # Import
```

## Hello to ggplot2

This uses the built-in `mtcars` dataset. To preview this dataset, you can use `summary(mtcars)` or `View(mtcars)`.

```{r}
View(mtcars)
```

We can see that there are the `mpg` (miles per gallons) and `wt` (weight) columns in the dataset. Let's plot the 2 dimensions on the scatterplot using `ggplot2`.

```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) + geom_point()
```

## Bar chart with iris dataset

We can also try making a bar chart with the built-in `iris` dataset.

```{r}
ggplot(data = iris, aes(x = Species, fill = Species)) + geom_bar()
```

## Customizing plots

Adding titles and labels:

```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  ggtitle("Scatter Plot of mpg vs wt") +
  xlab("Weight") +
  ylab("Miles Per Gallon")
```

Changing colors:

```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point()
```

### Experimenting with the Iris Dataset

-   **Dataset**: Iris (available in R by default)
-   **Task**: Create a scatter plot showing the relationship between petal length and petal width, colored by species.
-   **Customization**: Add a smooth regression line for each species.

```{r}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle("Petal Length vs Width by Species") +
    theme_minimal()
```

### Visualizing the mtcars Dataset

-   **Dataset**: mtcars (available in R by default)
-   **Task**: Create a bar plot showing the average miles per gallon (mpg) for cars with different numbers of cylinders.
-   **Customization**: Use a different fill color for each cylinder type and add labels for the average mpg.

```{r}
ggplot(mtcars, aes(x = factor(cyl), y = mpg, fill = factor(cyl))) +
    geom_bar(stat = "summary", fun = mean) +
    geom_text(stat = 'summary', aes(label = round(..y.., 1)), vjust = -0.5) +
    labs(x = "Number of Cylinders", y = "Average Miles per Gallon", title = "Average MPG by Cylinder Count") +
    theme_bw()
```

### Exploring the gapminder Dataset

-   **Dataset**: gapminder (install using **`install.packages("gapminder")`** and then **`library(gapminder)`**)
-   **Task**: Create a line plot showing GDP per capita over time for select countries.
-   **Customization**: Use different line types and colors for each country.

```{r}
# install.packages("gapminder")
library(gapminder)
ggplot(subset(gapminder, country %in% c("Japan", "United Kingdom", "United States")),
       aes(x = year, y = gdpPercap, color = country, linetype = country)) +
    geom_line() +
    scale_y_log10() +
    ggtitle("GDP Per Capita Over Time") +
    theme_light()
```

### Working with the diamonds Dataset

-   **Dataset**: diamonds (part of ggplot2 package)
-   **Task**: Create a histogram of diamond prices, faceted by cut quality.
-   **Customization**: Adjust the bin width and use a theme that enhances readability.

```{r}
ggplot(diamonds, aes(x = price)) +
    geom_histogram(binwidth = 500) +
    facet_wrap(~cut) +
    labs(title = "Diamond Prices by Cut Quality", x = "Price", y = "Count") +
    theme_classic()
```
